---
title: "Histograms"
subtitle: "Reading distributions with ggplot2"
format: live-html
engine: knitr
toc: true
toc-depth: 2
webr:
  packages:
    - ggplot2
---

{{< include _extensions/r-wasm/live/_knitr.qmd >}}

## Student Information

```{r}
#| label: student-information-1
#| echo: false
learnr2::student_info()
```

In this tutorial you will draw histograms of real data --- highway mileage
for 234 cars and prices for 53,940 diamonds --- learn to read their shape,
and write them up in a Quarto document of your own.

You work in two places. The code cells on this page run in your browser, so
you can experiment freely. The document you build lives in a GitHub repo
named `histograms`, which the next section creates by hand, with the real
commands rather than any shortcut script. By the end you will hand in the
URL of that repo, along with the answers this page saves as you go --- the
download button at the bottom collects them all.

## Connect to your repo

You should be doing this tutorial in a repo named `histograms`. You do not
have one yet. This section creates it --- by hand, with the same commands a
shortcut script would run for you --- so that every command is one you will
use again this term. Do **not** run `connect-repo` today.

Everything happens in the **bash Terminal** (Terminal → New Terminal, in
the Panel). Each question below shows what a correct run looks like. Paste
your own command and its response, then continue.

Open a new bash Terminal and run `pwd`, then `ls -a`.

```{r}
#| label: connect-to-your-repo-1
#| echo: false
learnr2::question("CP/CR.", type = "reflection_editable")
```

### What you should see

```
codespace-starter $ pwd
/workspaces/codespace-starter
codespace-starter $ ls -a
.  ..  .devcontainer  .github  .git
```

(Your listing may show a few extra files. That is fine.)

### Key idea

This is `codespace-starter` --- the workshop every course Codespace starts
from, not your project. Nothing you save here is yours; your work will live
in the repo you create in the next few steps.

Two commands make `gh` and `git` act as you instead of the Codespace.
Run `unset GITHUB_TOKEN GH_TOKEN`, then `gh auth status`. If it says you
are not logged in, sign in with
`gh auth login --hostname github.com --git-protocol https --web --clipboard=false`
(click Authorize in the browser that opens), then run `gh auth status`
again.

```{r}
#| label: connect-to-your-repo-2
#| echo: false
learnr2::question("CP/CR.", type = "reflection_editable")
```

### What you should see

```
codespace-starter $ gh auth status
github.com
  ✓ Logged in to github.com account your-username
  - Active account: true
  - Git operations protocol: https
  - Token: gho_************************************
  - Token scopes: 'gist', 'read:org', 'repo', 'workflow'
```

### Key idea

The `repo` scope is the one that matters --- it is what lets your login
create repositories and push to them. If you were already signed in, that
is a login from an earlier session; credentials stick around inside a
Codespace, but clearing the built-in token is what makes every later
command act as you.

Now make `git push` authenticate with your token --- from the bash Terminal
*and* from the Source Control panel. Run these four commands, one at a
time:

````
git config --global --replace-all credential.helper ""
git config --global --add credential.helper store
printf 'https://x-access-token:%s@github.com\n' "$(gh auth token)" > ~/.git-credentials
chmod 600 ~/.git-credentials
````

Then run `ls -l ~/.git-credentials`.

```{r}
#| label: connect-to-your-repo-3
#| echo: false
learnr2::question("CP/CR.", type = "reflection_editable")
```

### What you should see

```
codespace-starter $ ls -l ~/.git-credentials
-rw------- 1 rstudio rstudio 62 Oct  4 16:20 /home/rstudio/.git-credentials
```

### Key idea

That file holds your personal token --- the same thing a password would be.
The `-rw-------` permissions mean only you can read it, and git consults it
for every push, which is why the Source Control panel pushes as you rather
than as the Codespace.

Create this tutorial's repo. From the same bash Terminal, run:

````
cd /workspaces
gh repo create histograms --public --clone
````

Then run `ls`. (If git says the repository already exists, clone it
instead with `gh repo clone histograms`.)

```{r}
#| label: connect-to-your-repo-4
#| echo: false
learnr2::question("CP/CR.", type = "reflection_editable")
```

### What you should see

```
workspaces $ cd /workspaces
workspaces $ gh repo create histograms --public --clone
✓ Created repository your-username/histograms on github.com
  https://github.com/your-username/histograms
Cloning into 'histograms'...
warning: You appear to have cloned an empty repository.
workspaces $ ls
codespace-starter  histograms
```

(Your `ls` may show folders from earlier sessions. That is fine.)

### Key idea

One command did two jobs: it created the filing cabinet on GitHub, and
`--clone` copied it down to a local folder. The "empty repository" warning
is expected --- a brand-new repo has no commits yet. `--public` is what
later lets GitHub Pages host your work for free.

How does the local folder know which GitHub repo it belongs to? Run:

````
cd histograms
git remote -v
````

```{r}
#| label: connect-to-your-repo-5
#| echo: false
learnr2::question("CP/CR.", type = "reflection_editable")
```

### What you should see

```
workspaces $ cd histograms
histograms $ git remote -v
origin  https://github.com/your-username/histograms (fetch)
origin  https://github.com/your-username/histograms (push)
```

### Key idea

`origin` is git's nickname for the GitHub copy of this repo. From now on
`git push` sends your commits there and `git pull` brings changes back.

Point VS Code at your repo: File → Open Folder…, then open
`/workspaces/histograms`. The window reloads and this tutorial stops ---
that is expected. Restart it exactly as you started it: R Tutorials in
the Activity Bar, find this tutorial, click the arrow. Your saved answers
come back. Then open a new bash Terminal and run `pwd`.

```{r}
#| label: connect-to-your-repo-6
#| echo: false
learnr2::question("CP/CR.", type = "reflection_editable")
```

### What you should see

```
histograms $ pwd
/workspaces/histograms
```

### Key idea

New terminals now open in your own repo, and the Explorer, the Source
Control panel, and every AI agent all act on `/workspaces/histograms` ---
you are out of the workshop and into your own project.

## Your analysis workspace

Create `analysis.qmd` in your repo, with the Title Case title
*Histogram Analysis*. Give it a setup chunk holding `library(ggplot2)`
with `#| message: false`, and add `execute: echo: false` to the YAML, so
the rendered page shows results rather than code. Use AI however you like.
Then, in the bash Terminal, render it:

````
quarto render analysis.qmd
````

and right-click `analysis.html` in the Explorer → **Open with Live
Server**, keeping that tab open --- it refreshes on every later render.

```{r}
#| label: your-analysis-workspace-1
#| echo: false
learnr2::question("CP/CR.", type = "reflection_editable")
```

### What you should see

```
histograms $ quarto render analysis.qmd
processing file: analysis.qmd
output file: analysis.html
```

### Key idea

From now on, **Render** is shorthand for `quarto render analysis.qmd` in
the bash Terminal, then a glance at the Live Server tab. The `echo: false`
setting matters for a document someone else reads: they want the picture,
not the code that drew it.

Rendering created an `analysis_files` folder full of web assets git has
no business tracking. Add `analysis_files` to `.gitignore` (create the
file if there is none).

```{r}
#| label: your-analysis-workspace-2
#| echo: false
learnr2::question("Paste the `.gitignore` file.", type = "reflection_editable")
```

### What you should see

```
analysis_files
```

### Key idea

`.gitignore` is a list of files git should walk past. The code that
*produces* the renderable output belongs in the repo; the output itself
does not.

## Reading a histogram

A **histogram** cuts one numeric variable into bins and draws a bar whose
height is the count in each bin. It answers one question: what does the
distribution of this variable look like?

```{webr}
#| label: reading-a-histogram-1
#| autorun: true
library(ggplot2)
ggplot(mpg, aes(hwy)) +
  geom_histogram()
```

```{webr}
#| label: reading-a-histogram-3
#| exercise: ex_hwy
#| persist: true
library(ggplot2)

# Make a histogram of highway mileage (`hwy`) from `mpg`.
```

::: { .hint exercise="ex_hwy" }
`geom_histogram()` draws bars for a binned numeric variable; map `hwy` to
x inside `aes()`.
:::

::: { .solution exercise="ex_hwy" }
```r
library(ggplot2)
ggplot(mpg, aes(hwy)) +
  geom_histogram()
```
:::

```{r}
#| label: reading-a-histogram-2
#| echo: false
learnr2::question(
  "In a histogram, what does the height of each bar count?",
  learnr2::answer(
    "How many of the 234 cars fall in that bar's range of `hwy`",
    correct = TRUE
  ),
  learnr2::answer("How many distinct values of `hwy` exist"),
  learnr2::answer("The highway mileage of a single car"),
  learnr2::answer("The percentage of cars in that bar"),
  allow_retry = TRUE
)
```

### Key idea

Each bar covers a *range* of highway mileages --- a bin --- and its height
is the number of cars whose `hwy` falls inside that range. The 234 cars
are the `n` behind this plot, and a histogram never shows more than that.

```{r}
#| label: reading-a-histogram-4
#| echo: false
learnr2::question(
  "Run your cell and look at the plot. In one sentence, describe where the
  bulk of the 234 cars sit.",
  learnr2::answer(
    "Most cars get between roughly 18 and 30 mpg on the highway, with the
    tallest bars around 24 to 28, and a thin tail of a few cars reaching
    into the high 30s up to 44.",
    correct = TRUE
  ),
  type = "reflection"
)
```

### Key idea

Center, spread, shape, outliers --- those four words are the whole
vocabulary of a histogram. Here the center sits near 24 highway mpg (the
median), the values stretch about 12 to 44, and the tail runs to the right.

```{r}
#| label: reading-a-histogram-5
#| echo: false
learnr2::question(
  "Which of these belongs in a histogram?",
  learnr2::answer("Highway mileage of 234 cars", correct = TRUE),
  learnr2::answer("Counts of cars by drive type: fwd, 4wd, r"),
  learnr2::answer("Number of students in each letter grade"),
  learnr2::answer("Prices labeled cheap, medium, expensive"),
  allow_retry = TRUE
)
```

### Key idea

A histogram bins one *continuous* numeric variable; a bar chart counts the
levels of a *category*. The giveaway is the x-axis: a histogram's axis is a
number line, with no gaps between the bars.

## Choosing the bins

`geom_histogram()` draws 30 bins by default --- a starting point, not a
law. Change it with `bins =` (a count) or `binwidth =` (a width on the
data's own scale), and watch how the picture changes.

```{webr}
#| label: choosing-the-bins-1
#| autorun: true
library(ggplot2)
ggplot(mpg, aes(hwy)) +
  geom_histogram()
```

```{webr}
#| label: choosing-the-bins-3
#| exercise: ex_binwidth
#| persist: true
library(ggplot2)

# Redraw the `hwy` histogram with bins exactly 1 mpg wide.
```

::: { .hint exercise="ex_binwidth" }
Pass `binwidth = 1` to `geom_histogram()`.
:::

::: { .solution exercise="ex_binwidth" }
```r
library(ggplot2)
ggplot(mpg, aes(hwy)) +
  geom_histogram(binwidth = 1)
```
:::

```{r}
#| label: choosing-the-bins-2
#| echo: false
learnr2::question(
  "By default, how many bins does `geom_histogram()` draw?",
  learnr2::answer("30", correct = TRUE),
  learnr2::answer("30 bins", correct = TRUE),
  type = "text",
  allow_retry = TRUE
)
```

### Key idea

Thirty is ggplot2's default because it works okay for many datasets ---
not because statistics says 30 is correct. The number of bins is a display
choice you make, and you should say which choice you made.

```{r}
#| label: choosing-the-bins-4
#| echo: false
learnr2::question(
  "Run your cell. Compared with the 30-bin version, what changed about the
  shape?",
  learnr2::answer(
    "The smooth hump breaks into jagged spikes and gaps --- bins holding
    one or two cars now show up as isolated bars, so sampling noise reads
    as structure.",
    correct = TRUE
  ),
  type = "reflection"
)
```

### Key idea

Too few bins pile different shapes into the same bar and hide structure;
too many turn the luck of the draw into apparent spikes. Bin width is a
display choice --- the 234 cars never changed, only the ruler you drew
over them.

```{r}
#| label: choosing-the-bins-5
#| echo: false
learnr2::question(
  "You plot `hwy` twice, once with `bins = 5` and once with
  `bins = 100`, and the pictures look quite different. What can you
  honestly conclude?",
  learnr2::answer(
    "The bins are a display choice; the underlying 234 cars are identical
    in both pictures.",
    correct = TRUE
  ),
  learnr2::answer("The 100-bin plot is the more accurate one"),
  learnr2::answer("The 5-bin plot proves the distribution is symmetric"),
  learnr2::answer("Report whichever picture supports your story"),
  allow_retry = TRUE
)
```

### Key idea

Always report the bin width you used --- a reader who cannot reproduce your
picture cannot check your claim. When a feature appears at only one bin
width, treat it as a question, not a finding.

Now put it to work in your repo: add a histogram of `hwy` as the last
chunk of your `analysis.qmd`, Render, and look at it in Live Server.

```{r}
#| label: choosing-the-bins-6
#| echo: false
learnr2::question("Paste the chunk.", type = "reflection_editable")
```

### What you should see

```
ggplot(mpg, aes(hwy)) +
  geom_histogram()
```

### Key idea

Your working chunk starts simple and gets rewritten as you learn more ---
each exercise replaces what was there before, so only the final state
matters. Nothing is lost by changing your mind; the history lives in git.

Your first histogram is in. In the bash Terminal, save it:

````
git add analysis.qmd && git commit -m "Add hwy histogram" && git push -u origin main
````

```{r}
#| label: choosing-the-bins-7
#| echo: false
learnr2::question("CP/CR.", type = "reflection_editable")
```

### What you should see

```
histograms $ git add analysis.qmd && git commit -m "Add hwy histogram" && git push -u origin main
[main (root-commit) a1b2c3d] Add hwy histogram
 1 file changed, 42 insertions(+)
Enumerating objects: 5, done.
Counting objects: 100% (5/5), done.
To https://github.com/your-username/histograms.git
 * [new branch]      main -> main
branch 'main' set up to track 'origin/main'.
```

### Key idea

The first push needs `-u origin main`: your local `main` branch has never
met `origin`, so git has to pair them; after that, plain `git push` is
enough. And push as you commit --- a Codespace is disposable, so unpushed
work is the only work you can lose.

## Center, spread, and shape

Diamonds next: 53,940 of them, with prices from $44 to $18,823. Before
plotting, look at the numbers.

```{webr}
#| label: center-spread-and-shape-1
#| autorun: true
library(ggplot2)
summary(diamonds$price)
```

```{webr}
#| label: center-spread-and-shape-3
#| exercise: ex_price
#| persist: true
library(ggplot2)

# Histogram of `price` from `diamonds`, with one vertical line at the mean
# and another at the median.
```

::: { .hint exercise="ex_price" }
Add two `geom_vline()` layers, each with `xintercept = mean(diamonds$price)`
or `xintercept = median(diamonds$price)`, and give them different colors.
:::

::: { .solution exercise="ex_price" }
```r
library(ggplot2)
ggplot(diamonds, aes(price)) +
  geom_histogram() +
  geom_vline(xintercept = mean(diamonds$price), color = "red") +
  geom_vline(xintercept = median(diamonds$price), color = "blue")
```
:::

```{r}
#| label: center-spread-and-shape-2
#| echo: false
learnr2::question(
  "Which two statements about the shape of `price` are true?",
  learnr2::answer(
    "The mean ($3,933) sits well above the median ($2,401)",
    correct = TRUE
  ),
  learnr2::answer("The distribution is skewed to the right", correct = TRUE),
  learnr2::answer("The distribution is symmetric about the mean"),
  learnr2::answer("The median sits above the third quartile"),
  allow_retry = TRUE
)
```

### Key idea

When mean and median disagree, the tail wins: a long right tail drags the
mean upward, while the median stays at the 50% mark. Skew is named for the
tail, never the hump.

```{r}
#| label: center-spread-and-shape-4
#| echo: false
learnr2::question(
  "For this distribution, which is larger: the mean or the median?",
  learnr2::answer("mean", correct = TRUE),
  learnr2::answer("the mean", correct = TRUE),
  type = "text",
  allow_retry = TRUE
)
```

### Key idea

The mean line sits to the right of the median because a few very expensive
diamonds pull the average up. When a distribution is skewed, the median is
the safer summary of "typical" --- it cannot be moved by a handful of
extreme values.

Rewrite your working chunk: a histogram of `price` with the mean and
median lines, axis labels, and the number of diamonds in the title
(n = 53,940). Render.

```{r}
#| label: center-spread-and-shape-5
#| echo: false
learnr2::question("Paste the chunk.", type = "reflection_editable")
```

### What you should see

```
ggplot(diamonds, aes(price)) +
  geom_histogram() +
  geom_vline(xintercept = mean(diamonds$price), color = "red") +
  geom_vline(xintercept = median(diamonds$price), color = "blue") +
  labs(
    title = "Prices of 53,940 diamonds",
    x = "Price (USD)",
    y = "Number of diamonds"
  )
```

### Key idea

Every histogram should say how many observations it counts --- filters,
missing values, and group definitions can silently shrink `n`. The title is
the cheapest place to say it, and "Number of diamonds" beats a bare "count"
for telling a reader what the bars hold.

```{r}
#| label: center-spread-and-shape-6
#| echo: false
learnr2::question(
  "Look at your rendered plot and write one sentence interpreting it: what
  would a buyer take away? Submit, then compare with our answer.",
  learnr2::answer(
    "Half of the 53,940 diamonds are priced under about $2,400, but a long
    tail of expensive stones pulls the average up to nearly $3,900 --- so a
    typical diamond costs closer to the median than to the mean.",
    correct = TRUE
  ),
  type = "reflection"
)
```

### Key idea

An interpretation names the group, the feature, and the comparison ---
"most …, but …" --- and stays inside what the plot can support. Notice
what this one cannot say: nothing about *why* prices are skewed.

## Comparing groups

```{webr}
#| label: comparing-groups-1
#| exercise: ex_facet
#| persist: true
library(ggplot2)

# Histogram of `hwy`, one panel per `class` of car.
```

::: { .hint exercise="ex_facet" }
`facet_wrap(~ class)` splits one plot into a panel per level of `class`.
:::

::: { .solution exercise="ex_facet" }
```r
library(ggplot2)
ggplot(mpg, aes(hwy)) +
  geom_histogram() +
  facet_wrap(~ class)
```
:::

```{r}
#| label: comparing-groups-2
#| echo: false
learnr2::question(
  "Run your cell. Which classes sit clearly at the bottom of highway
  mileage?",
  learnr2::answer("pickup and suv", correct = TRUE),
  learnr2::answer("minivan and suv"),
  learnr2::answer("pickup and 2seater"),
  learnr2::answer("midsize and minivan"),
  allow_retry = TRUE
)
```

### Key idea

Pickups (about 17 mpg) and SUVs (about 18) sit far to the left of compact
and subcompact cars (about 28). Facets share one x-axis, so the panels are
directly comparable --- but each panel counts only its own group (62 SUVs
but just 5 two-seaters), so compare shapes, not bar heights.

Add the faceted histogram to `analysis.qmd` as your working chunk ---
labels and `n` in the title included. Render.

```{r}
#| label: comparing-groups-3
#| echo: false
learnr2::question("Paste the chunk.", type = "reflection_editable")
```

### What you should see

```
ggplot(mpg, aes(hwy)) +
  geom_histogram() +
  facet_wrap(~ class) +
  labs(
    title = "Highway mileage by class (234 cars)",
    x = "Highway miles per gallon",
    y = "Number of cars"
  )
```

### Key idea

Seven panels means seven group sizes: the two-seater panel holds just 5
cars, so its shape is fragile, while the SUV panel's is dependable.
Faceting answers "how do the distributions differ?" --- the question a
table of averages cannot.

Commit this section's progress in the bash Terminal:

````
git add analysis.qmd && git commit -m "Add faceted histogram" && git push
````

```{r}
#| label: comparing-groups-4
#| echo: false
learnr2::question("CP/CR.", type = "reflection_editable")
```

### What you should see

```
histograms $ git add analysis.qmd && git commit -m "Add faceted histogram" && git push
[main e5f6a7b] Add faceted histogram
 1 file changed, 16 insertions(+), 6 deletions(-)
To https://github.com/your-username/histograms.git
   a1b2c3d..e5f6a7b  main -> main
```

### Key idea

Notice the push now moves `a1b2c3d..e5f6a7b` --- each commit points back
at the one before it, so GitHub holds a history, not just the latest file.
Plain `git push` will do from now on; `-u` was a first-push-only
courtesy.

## Transforming the scale

```{webr}
#| label: transforming-the-scale-1
#| exercise: ex_log
#| persist: true
library(ggplot2)

# Histogram of `price` on a log10 scale.
```

::: { .hint exercise="ex_log" }
Add `scale_x_log10()` after the histogram layer to change the x-axis
scale without touching the data.
:::

::: { .solution exercise="ex_log" }
```r
library(ggplot2)
ggplot(diamonds, aes(price)) +
  geom_histogram() +
  scale_x_log10()
```
:::

```{r}
#| label: transforming-the-scale-2
#| echo: false
learnr2::question(
  "Run your cell. In one sentence, what does the log scale do to the
  shape?",
  learnr2::answer(
    "The right-skewed hump becomes roughly symmetric and centered --- equal
    distances on the axis now mean equal ratios, so each tick is ten
    times the last.",
    correct = TRUE
  ),
  type = "reflection"
)
```

### Key idea

A log axis trades "how much more" for "how many times more". Left of
center are the cheap diamonds, right are the expensive ones, and the hump
shows where most prices actually cluster.

```{r}
#| label: transforming-the-scale-3
#| echo: false
learnr2::question(
  "Why log-transform a right-skewed variable like `price` before comparing
  groups or reporting a center?",
  learnr2::answer(
    "Because a multiplicative difference becomes a fixed distance on the
    log scale, so spread and center become comparable across groups.",
    correct = TRUE
  ),
  learnr2::answer("Because it makes the sample size effectively larger"),
  learnr2::answer("Because logs delete outliers, which are usually errors"),
  learnr2::answer("Because the mean of logged values equals the logged mean"),
  allow_retry = TRUE
)
```

### Key idea

Log transformation does not delete anything and does not change the data:
it changes the ruler. On a log scale, $1,000 to $10,000 is the same
distance as $100 to $1,000, because both are a tenfold jump.

Swap the log-scale version into `analysis.qmd` as your working chunk ---
keep the labels and `n` in the title. Render.

```{r}
#| label: transforming-the-scale-4
#| echo: false
learnr2::question("Paste the chunk.", type = "reflection_editable")
```

### What you should see

```
ggplot(diamonds, aes(price)) +
  geom_histogram() +
  scale_x_log10() +
  labs(
    title = "Prices of 53,940 diamonds (log10 scale)",
    x = "Price (USD, log10 scale)",
    y = "Number of diamonds"
  )
```

### Key idea

Say "log scale" in the label itself --- an unlabeled log axis invites
readers to misread $3,000 as $3. Here the ticks run $100, $1,000,
$10,000, and each step is a multiplication.

## Caching

Each render redraws every histogram, including the ones built from all
53,940 diamonds. Quarto can remember a chunk's results and skip re-running
it --- **caching** --- which you set up in three steps.

Step one, clean up. Find the chunk you want to cache --- the one that
draws the diamond price histogram --- and delete any exploratory lines
you left in it (test prints, scratch summaries), so it holds only the
final plot.

```{r}
#| label: caching-1
#| echo: false
learnr2::question("Paste the chunk.", type = "reflection_editable")
```

### What you should see

```
ggplot(diamonds, aes(price)) +
  geom_histogram() +
  scale_x_log10() +
  labs(
    title = "Prices of 53,940 diamonds (log10 scale)",
    x = "Price (USD, log10 scale)",
    y = "Number of diamonds"
  )
```

### Key idea

A cached chunk should hold a definition and nothing else --- anything you
add later invalidates the cache and forces the slow work all over again.

Step two, turn it on. Add `#| cache: true` as that chunk's first line,
Render, then run `ls` in the bash Terminal.

```{r}
#| label: caching-2
#| echo: false
learnr2::question("CP/CR.", type = "reflection_editable")
```

### What you should see

```
histograms $ ls
analysis.html  analysis.qmd  analysis_cache  analysis_files
```

### Key idea

`analysis_cache/` appearing next to your document is the visible evidence
that caching is wired up. Quarto now reuses that chunk's results until you
edit its code.

Step three, keep git clean. The Source Control change count just jumped
because `analysis_cache/` filled with files git should not track. Add
`analysis_cache` to `.gitignore`.

```{r}
#| label: caching-3
#| echo: false
learnr2::question("Paste the `.gitignore` file.", type = "reflection_editable")
```

### What you should see

```
analysis_files
analysis_cache
```

### Key idea

A cache is derived --- git tracks what produces it (the code), not the
products. Add each cache directory once, and the change count drops back.

## Summary

Final check of the finished document: Render, open `analysis.html` with
Live Server, and take a screenshot of the page --- `Cmd+Shift+4` on a
Mac, `Win+Shift+S` on Windows, `Ctrl+Show windows` on a Chromebook.
Submit, then paste the screenshot into the box (Ctrl/Cmd+V).

```{r}
#| label: summary-1
#| echo: false
learnr2::question(
  "Paste the screenshot.",
  learnr2::answer(
    "A rendered page of analysis.qmd showing four labeled histograms of
    real data: hwy, price with mean and median lines, hwy faceted by
    class, and price on a log10 scale.",
    correct = TRUE
  ),
  type = "reflection_editable",
  allow_image = TRUE
)
```

### Key idea

One page, four histograms, each labeled with what it counts and how many
observations back it --- that is an artifact you would show someone, not a
worksheet you would hand in and forget.

Commit any remaining changes:

````
git add -A && git commit -m "Finish histograms tutorial" && git push
````

(if git reports nothing to commit, you are already up to date). Then show
your repo's timeline:

````
git log --oneline
````

```{r}
#| label: summary-2
#| echo: false
learnr2::question("CP/CR.", type = "reflection_editable")
```

### What you should see

```
histograms $ git log --oneline
f7a8b9c Finish histograms tutorial
e5f6a7b Add faceted histogram
a1b2c3d Add hwy histogram
```

(Your hashes --- the short codes at the start of each line --- will differ
from ours. That is expected: they name your commits, not ours.)

### Key idea

Each commit points back at the one before it, so `git log` is a timeline
of your work rather than a stack of file copies. Push every checkpoint, so
the timeline lives on GitHub rather than inside a disposable Codespace.

Finally, open your repo on GitHub --- the page someone would open to see
your work.

```{r}
#| label: summary-3
#| echo: false
learnr2::question(
  "Paste the URL of your GitHub repo.",
  learnr2::answer(
    "https://github.com/your-username/histograms --- yours will carry your
    own GitHub username in place of `your-username`.",
    correct = TRUE
  ),
  type = "reflection_editable"
)
```

### Key idea

That URL is your submission: your code and your full history live there,
pushed from a Codespace that could be deleted tomorrow. The download button
below collects every answer you gave in this tutorial --- turn that file in
as well.

## Your Answers

```{r}
#| label: your-answers-1
#| echo: false
learnr2::question(
  "How many minutes, approximately, did it take you to complete this
  tutorial? For example, an hour and a half would be 90 minutes.",
  type = "reflection_editable",
  validate = "integer"
)
```

```{r}
#| label: your-answers-2
#| echo: false
learnr2::download_answers_button(filename_prefix = "06-histograms")
```
