---
title: "Linear Regression with Multiple Predictors"
subtitle: "Indicator and categorical predictors, many predictors at once, adjusted R-squared, and stepwise model selection."
format: live-html
engine: knitr
toc: true
toc-depth: 2
learnr2:
  ordering: 8
---

{{< include _extensions/r-wasm/live/_knitr.qmd >}}

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(openintro)
library(broom)
knitr::opts_chunk$set(out.width = "90%")
```

## Student Information

```{r}
#| label: student-information-1
#| echo: false
learnr2::student_info()
```

## Introduction

This tutorial covers [Chapter 8: Linear regression with multiple predictors](https://openintrostat.github.io/ims/model-mlr.html) from [*Introduction to Modern Statistics*](https://openintrostat.github.io/ims/) by Mine Çetinkaya-Rundel and Johanna Hardin. You will learn how a regression model uses indicator and categorical predictors, how to fit and read a model with many predictors at once, and how to choose among models with adjusted R-squared, using backward elimination and forward selection.

### Exercise 1

You should be doing this tutorial in a repo named `linear-regression-with-multiple-predictors`. If you are not, create one and then connect to it. You may need to restart this tutorial after you do so.

Create a new file, `analysis.qmd`, with the title `"Linear Regression with Multiple Predictors"` and your name as the author. In a bash Terminal, render it:

```
quarto render analysis.qmd
```

Open `analysis.html` with Live Server (right-click it in the Explorer → **Open with Live Server**) and keep the tab open. It refreshes on every render. Going forward, we will just tell you to "Render" when we want you to take these steps.

Create a `.gitignore` with `analysis_files` followed by a blank line. Commit and push. Use AI however you like.

If `show_file()` fails, it is probably because you have not yet loaded `library(learnr2)` in the R Terminal.

```{r}
#| label: introduction-1
#| echo: false
learnr2::question('In the R Terminal, run show_file(".gitignore"). CP/CR.', type = "reflection")
```

###

```
analysis_files
```

###

"Multiple regression extends single predictor variable regression to the case that still has one response but many predictors." Our one response is the interest rate on a loan, and our predictors describe the loan and the borrower.

### Exercise 2

Add a chunk to `analysis.qmd` that loads the **[tidyverse](https://www.tidyverse.org/)**, with `#| message: false` so the startup messages stay out of the rendered page. Add `execute: echo: false` to the YAML header. Render.

```{r}
#| label: introduction-2
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

<pre><code>#| message: false
library(tidyverse)
</code></pre>

###

Base R fits linear models without any extra package. The **tidyverse** does the rest of the work here: preparing the data, looping over many candidate models, and plotting the results.

### Exercise 3

Add **[openintro](https://openintrostat.github.io/openintro/)** to that chunk. Render.

```{r}
#| label: introduction-3
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

<pre><code>#| message: false
library(tidyverse)
library(openintro)
</code></pre>

###

"The `loans_full_schema` data can be found in the **openintro** R package." You met it in the Exploring Categorical Data tutorial; this time we use it to model interest rates.

### Exercise 4

Add **[broom](https://broom.tidymodels.org/)** to that chunk. Render.

```{r}
#| label: introduction-4
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

<pre><code>#| message: false
library(tidyverse)
library(openintro)
library(broom)
</code></pre>

###

**broom** turns a fitted model into tibbles: one row per coefficient, or one row per model holding its R-squared and adjusted R-squared. It belongs to the **[tidymodels](https://www.tidymodels.org/)** family, and it is what lets us compare dozens of models in a single table.

## Categorical and many predictors

"The outcome variable we would like to better understand is the interest rate assigned to the loan." We build the chapter's `loans` data, then fit models with an indicator predictor, a categorical predictor, and finally every predictor at once.

### Exercise 1

Add a chunk that prints `loans_full_schema`. Render.

```{r}
#| label: categorical-and-many-predictors-1
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-1-test
#| echo: true
loans_full_schema
```

###

10,000 loans and 55 variables, far more than the chapter needs. "The loan data includes terms of the loan as well as information about the borrower."

### Exercise 2

```{r}
#| label: categorical-and-many-predictors-2
#| echo: false
learnr2::question("In the R Terminal, run ?loans_full_schema. Copy and paste the entries for inquiries_last_12m, total_credit_limit, total_credit_utilized, and public_record_bankrupt from the Format section.", type = "reflection")
```

###

```
     inquiries_last_12m Inquiries into the applicant's credit during
          the last 12 months.

     total_credit_limit Total available credit, e.g. if only credit
          cards, then the total of all the credit limits. This excludes
          a mortgage.

     total_credit_utilized Total credit balance, excluding a mortgage.

     public_record_bankrupt Number of bankruptcies listed in the public
          record for this applicant.
```

###

The chapter builds two new variables from these: "`credit_util` which is calculated as the total credit utilized divided by the total credit limit and `bankruptcy` which turns the number of bankruptcies to an indicator variable (0 for no bankruptcies and 1 for at least 1 bankruptcy)." It also renames `inquiries_last_12m` to `credit_checks`.

### Exercise 3

Replace the code in that chunk so that it counts the loans at each level of `verified_income`, including any level that has no loans. Render.

```{r}
#| label: categorical-and-many-predictors-3
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-3-test
#| echo: true
loans_full_schema |>
  count(verified_income, .drop = FALSE)
```

###

The factor carries a fourth level, an empty string, that not one loan uses. The chapter drops that unused level, so `verified_income` has three: Not Verified, Source Verified, and Verified.

### Exercise 4

Replace the code in that chunk so that it counts the loans with each value of `public_record_bankrupt`. Render.

```{r}
#| label: categorical-and-many-predictors-4
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-4-test
#| echo: true
loans_full_schema |>
  count(public_record_bankrupt)
```

###

8,785 borrowers have no bankruptcy and only 20 have more than one, so a yes-or-no version loses little. "Using an indicator variable in place of a category name allows for these variables to be directly used in regression."

### Exercise 5

Start a new chunk that creates `loans` from `loans_full_schema`, keeping only `interest_rate`, `verified_income`, `debt_to_income`, `term`, `issue_month`, and `inquiries_last_12m`, renamed to `credit_checks`. Print `loans`. Render.

```{r}
#| label: categorical-and-many-predictors-5
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-5-test
#| echo: true
loans <- loans_full_schema |>
  rename(credit_checks = inquiries_last_12m) |>
  select(interest_rate, verified_income, debt_to_income,
         term, credit_checks, issue_month)

loans
```

###

Every loan is either 36 or 60 months long. For `issue_month`, the chapter's table says: "The month and year the loan was issued, which for these loans is always during the first quarter of 2018."

### Exercise 6

Add a new variable to `loans`, `credit_util`: the total credit utilized divided by the total credit limit. Keep it, placed after `debt_to_income`. Render.

```{r}
#| label: categorical-and-many-predictors-6
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-6-test
#| echo: true
loans <- loans_full_schema |>
  mutate(credit_util = total_credit_utilized / total_credit_limit) |>
  rename(credit_checks = inquiries_last_12m) |>
  select(interest_rate, verified_income, debt_to_income, credit_util,
         term, credit_checks, issue_month)

loans
```

###

`credit_util` answers the question, "Of all the credit available to the borrower, what fraction are they utilizing." A value of 0.548 means the first borrower is using about 55% of her available credit.

### Exercise 7

Add two more changes to `loans`. Create `bankruptcy`, a factor that is 1 if `public_record_bankrupt` is at least 1 and 0 otherwise, placed after `credit_util`. Drop the unused level of `verified_income`. Render.

```{r}
#| label: categorical-and-many-predictors-7
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-7-test
#| echo: true
loans <- loans_full_schema |>
  mutate(credit_util = total_credit_utilized / total_credit_limit,
         bankruptcy = as.factor(if_else(public_record_bankrupt == 0, 0, 1)),
         verified_income = droplevels(verified_income)) |>
  rename(credit_checks = inquiries_last_12m) |>
  select(interest_rate, verified_income, debt_to_income, credit_util,
         bankruptcy, term, credit_checks, issue_month)

loans
```

###

This is the chapter's `loans`, and every model in this tutorial is built from it. If your columns or their types differ from ours, replace your code with ours before going on.

### Exercise 8

Replace the line that prints `loans` with a summary of every variable in `loans`. Render.

```{r}
#| label: categorical-and-many-predictors-8
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-8-test
#| echo: true
loans <- loans_full_schema |>
  mutate(credit_util = total_credit_utilized / total_credit_limit,
         bankruptcy = as.factor(if_else(public_record_bankrupt == 0, 0, 1)),
         verified_income = droplevels(verified_income)) |>
  rename(credit_checks = inquiries_last_12m) |>
  select(interest_rate, verified_income, debt_to_income, credit_util,
         bankruptcy, term, credit_checks, issue_month)

summary(loans)
```

###

`debt_to_income` is missing for 24 loans, all of them joint applications, and `credit_util` for 2 borrowers whose credit limit is zero. `issue_month` still carries an empty level with no loans, which the chapter left alone.

### Exercise 9

Delete the summary, so that the chunk only creates `loans`. Render.

```{r}
#| label: categorical-and-many-predictors-9
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

<pre><code>loans <- loans_full_schema |>
  mutate(credit_util = total_credit_utilized / total_credit_limit,
         bankruptcy = as.factor(if_else(public_record_bankrupt == 0, 0, 1)),
         verified_income = droplevels(verified_income)) |>
  rename(credit_checks = inquiries_last_12m) |>
  select(interest_rate, verified_income, debt_to_income, credit_util,
         bankruptcy, term, credit_checks, issue_month)
</code></pre>

###

"Notice that the past bankruptcy variable (`bankruptcy`) is an indicator variable, where it takes the value 1 if the borrower had a past bankruptcy in their record and 0 if not." `verified_income` and `issue_month` are the other two categorical variables.

### Exercise 10

Add `#| cache: true` to that chunk. Render.

```{r}
#| label: categorical-and-many-predictors-10
#| echo: false
learnr2::question("In the bash Terminal, run ls. CP/CR.", type = "reflection")
```

###

```
analysis.html  analysis.qmd  analysis_cache  analysis_files
```

###

`analysis_cache` now holds `loans`, so later renders load it from disk. A cached chunk only re-runs when its code changes.

### Exercise 11

Your Source Control count just jumped, because Git sees every file in the new cache directory. Add `analysis_cache` to `.gitignore`, and the count drops back.

```{r}
#| label: categorical-and-many-predictors-11
#| echo: false
learnr2::question('In the R Terminal, run show_file(".gitignore"). CP/CR.', type = "reflection")
```

###

```
analysis_files
analysis_cache
```

###

The cache is a copy of results that anyone can rebuild from `analysis.qmd`, so it does not belong on GitHub. Only the code that made it does.

### Exercise 12

Start a new chunk that calculates the mean `interest_rate` and the number of loans for each value of `bankruptcy`. Render.

```{r}
#| label: categorical-and-many-predictors-12
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-12-test
#| echo: true
loans |>
  summarize(mean_rate = mean(interest_rate), n = n(), .by = bankruptcy)
```

###

Borrowers with a bankruptcy pay 13.1% on average, against 12.3% for the 8,785 without one. A regression with `bankruptcy` as its only predictor turns that gap into a slope.

### Exercise 13

Replace the code in that chunk so that it fits a linear model of `interest_rate` with `bankruptcy` as the only predictor, and shows just the term and the estimate for each coefficient, to two decimal places. Render.

```{r}
#| label: categorical-and-many-predictors-13
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-13-test
#| echo: true
lm(interest_rate ~ bankruptcy, data = loans) |>
  tidy() |>
  select(term, estimate) |>
  mutate(estimate = num(estimate, digits = 2))
```

###

The intercept, 12.34, is the mean rate without a bankruptcy, and 12.34 + 0.74 is the mean with one. "A slope of 0.74 means that the model predicts a 0.74% higher interest rate for those borrowers with a bankruptcy in their record."

### Exercise 14

Replace the predictor in that model with `verified_income`. Render.

```{r}
#| label: categorical-and-many-predictors-14
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-14-test
#| echo: true
lm(interest_rate ~ verified_income, data = loans) |>
  tidy() |>
  select(term, estimate) |>
  mutate(estimate = num(estimate, digits = 2))
```

###

Three levels, but only two coefficients: Not Verified is missing. "The missing level is called the reference level and it represents the default level that other levels are measured against."

### Exercise 15

Replace the code in that chunk so that it calculates the mean `interest_rate` for each level of `verified_income`. Render.

```{r}
#| label: categorical-and-many-predictors-15
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-15-test
#| echo: true
loans |>
  summarize(mean_rate = mean(interest_rate), .by = verified_income)
```

###

Not Verified averages 11.1%, the intercept; Source Verified averages 11.10 + 1.42 = 12.5%; and Verified 11.10 + 3.25 = 14.35%. "Intuitively, we would think that a loan would look *less* risky if the borrower's income has been verified," but "perhaps lenders require borrowers with poor credit to verify their income."

### Exercise 16

Replace the code in that chunk so that it fits a linear model of `interest_rate` using every other variable in `loans` as a predictor, and shows just the term and the estimate for each coefficient, to two decimal places. Render.

```{r}
#| label: categorical-and-many-predictors-16
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-16-test
#| echo: true
lm(interest_rate ~ ., data = loans) |>
  tidy() |>
  select(term, estimate) |>
  mutate(estimate = num(estimate, digits = 2))
```

###

Seven predictors produce nine slopes, because "a categorical predictor with $p$ different levels will be represented by $p - 1$ terms in a multiple regression model." The `bankruptcy` slope fell from 0.74 to 0.39: the one-predictor model "was constructed in a vacuum and did not consider the full context of everything that is considered when an interest rate is decided."

### Exercise 17

Replace the code in that chunk so that it shows the number of loans the full model actually used. Render.

```{r}
#| label: categorical-and-many-predictors-17
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-17-test
#| echo: true
lm(interest_rate ~ ., data = loans) |>
  nobs()
```

###

The chapter says "10,000 residuals are calculated, one for each observation," but the model used 9,974 loans. The 26 loans missing `debt_to_income` or `credit_util` were silently dropped, which is why it pays to check what a model was actually fit to.

### Exercise 18

Replace the code in that chunk with a linear model of `interest_rate` using only `credit_util` and `verified_income` as predictors, showing just the term and the estimate for each coefficient, to two decimal places. Render.

```{r}
#| label: categorical-and-many-predictors-18
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-18-test
#| echo: true
lm(interest_rate ~ credit_util + verified_income, data = loans) |>
  tidy() |>
  select(term, estimate) |>
  mutate(estimate = num(estimate, digits = 2))
```

###

The model gives every level of `verified_income` the same slope for `credit_util`, 4.80, and a different intercept. "Each coefficient represents how the model predicts the outcome might change with one unit increase of that predictor *given* the rest of the predictor variables in the model."

### Exercise 19

Replace the code in that chunk with a scatterplot of `interest_rate` against `credit_util`, colored by `verified_income`, with the fitted lines from that two-predictor model drawn on top. Use only the loans the model was fit to. Render.

```{r}
#| label: categorical-and-many-predictors-19
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-19-test
#| echo: true
lm(interest_rate ~ credit_util + verified_income, data = loans) |>
  augment() |>
  ggplot(aes(x = credit_util, color = verified_income)) +
  geom_point(aes(y = interest_rate), alpha = 0.1) +
  geom_line(aes(y = .fitted))
```

###

Three parallel lines: the categorical predictor shifts the line up or down, and `credit_util` sets its slope. A few points sit to the right of 1: 154 borrowers owe more than their credit limit.

### Exercise 20

Improve the scatterplot. Show credit utilization as a percentage, and add a title, axis labels, a legend title, a subtitle that states the main finding, and a caption giving the number of loans and the source. Render.

```{r}
#| label: categorical-and-many-predictors-20
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: categorical-and-many-predictors-20-test
#| echo: true
lm(interest_rate ~ credit_util + verified_income, data = loans) |>
  augment() |>
  ggplot(aes(x = credit_util, color = verified_income)) +
  geom_point(aes(y = interest_rate), alpha = 0.1) +
  geom_line(aes(y = .fitted), linewidth = 1) +
  scale_x_continuous(labels = scales::label_percent()) +
  labs(title = "Interest Rate by Credit Utilization and Income Verification",
       subtitle = "At every utilization, verified borrowers are predicted to pay 3.3 points more",
       x = "Credit Utilization",
       y = "Interest Rate (%)",
       color = "Income Verification",
       caption = "9,998 Lending Club loans, January to March 2018; openintro package") +
  theme(legend.position = "bottom")
```

###

"While we remain cautious about making any causal interpretations using multiple regression on observational data, such models are a common first step in gaining insights or providing some evidence of a causal connection." Here, verification may mark borrowers the lender already worried about, rather than cause their higher rates.

### Exercise 21

Commit `analysis.qmd` and `.gitignore` with the message `"Fit models with categorical and many predictors"` and push.

```{r}
#| label: categorical-and-many-predictors-21
#| echo: false
learnr2::question("In the bash Terminal, run git log --oneline -3. CP/CR.", type = "reflection")
```

###

```
3f8a1c2 (HEAD -> main, origin/main) Fit models with categorical and many predictors
9be04d2 Initial commit
```

###

Predictors that travel together are a problem for interpretation. "We say the two predictor variables are collinear (pronounced as *co-linear*) when they are correlated, and this multicollinearity complicates model estimation."

## Adjusted R-squared and model selection

"One challenge to working with multiple variables is that it is sometimes difficult to know which variables are most important to include in the model." The chapter's tool for that choice is adjusted R-squared, and its two strategies are backward elimination and forward selection.

### Exercise 1

Start a new chunk that fits the full model of `interest_rate` on every other variable in `loans` and shows its R-squared and adjusted R-squared, to four decimal places. Render.

```{r}
#| label: adjusted-r-squared-and-model-selection-1
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: adjusted-r-squared-and-model-selection-1-test
#| echo: true
lm(interest_rate ~ ., data = loans) |>
  glance() |>
  select(r.squared, adj.r.squared) |>
  mutate(across(everything(), \(x) num(x, digits = 4)))
```

###

The full model explains about 26% of the variability in interest rates, and the adjusted value, 0.2597, is a hair smaller. "The regular $R^2$ is a biased estimate of the amount of variability explained by the model when applied to model with more than one predictor."

### Exercise 2

Replace the code in that chunk so that it calculates both numbers by hand from the full model's residuals and the outcomes of the loans it used: the variance of the residuals, the variance of `interest_rate`, the number of loans `n`, R-squared, and adjusted R-squared with `k = 9` predictors, all to four decimal places. Render.

```{r}
#| label: adjusted-r-squared-and-model-selection-2
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: adjusted-r-squared-and-model-selection-2-test
#| echo: true
lm(interest_rate ~ ., data = loans) |>
  augment() |>
  summarize(var_resid = var(.resid),
            var_outcome = var(interest_rate),
            n = n()) |>
  mutate(r_squared = 1 - var_resid / var_outcome,
         adj_r_squared = 1 - var_resid / var_outcome * (n - 1) / (n - 9 - 1)) |>
  mutate(across(-n, \(x) num(x, digits = 4)))
```

###

The hand calculation matches the software. The adjusted version multiplies by $(n - 1) / (n - k - 1)$, and "because $k$ is never negative, the adjusted $R^2$ will be smaller -- often times just a little smaller -- than the unadjusted $R^2$."

### Exercise 3

Replace the code in that chunk so that it adds a column of pure random noise to `loans`, fits the full model with that noise as one more predictor, and shows R-squared and adjusted R-squared to five decimal places. Set the random seed to 8 first. Render.

```{r}
#| label: adjusted-r-squared-and-model-selection-3
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: adjusted-r-squared-and-model-selection-3-test
#| echo: true
set.seed(8)
loans |>
  mutate(noise = rnorm(n())) |>
  lm(interest_rate ~ ., data = _) |>
  glance() |>
  select(r.squared, adj.r.squared) |>
  mutate(across(everything(), \(x) num(x, digits = 5)))
```

###

Noise nudged R-squared up, from 0.26036 to 0.26040, while adjusted R-squared fell, from 0.25969 to 0.25966. "If we were to make predictions for *new data* using our current model, we would find that the unadjusted $R^2$ would tend to be slightly overly optimistic, while the adjusted $R^2$ formula helps correct this bias."

### Exercise 4

Replace the code in that chunk so that, for each of the seven predictors in turn, it fits the full model without that one predictor and records its adjusted R-squared. Show the results in a table sorted from highest to lowest, with adjusted R-squared to four decimal places. Render.

```{r}
#| label: adjusted-r-squared-and-model-selection-4
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: adjusted-r-squared-and-model-selection-4-test
#| echo: true
tibble(dropped = c("verified_income", "debt_to_income", "credit_util",
                   "bankruptcy", "term", "credit_checks", "issue_month")) |>
  mutate(adj_r_squared = map_dbl(dropped, \(x) {
    lm(interest_rate ~ ., data = select(loans, -all_of(x))) |>
      glance() |>
      pull(adj.r.squared)
  })) |>
  arrange(desc(adj_r_squared)) |>
  mutate(adj_r_squared = num(adj_r_squared, digits = 4))
```

###

This is the first step of **backward elimination**, which "starts with the full model -- the model that includes all potential predictor variables." Dropping `issue_month` raises adjusted R-squared from 0.2597 to 0.2598, so it goes; dropping `term` costs the most.

### Exercise 5

Replace the code in that chunk so that it repeats the step with `issue_month` already removed: for each of the six remaining predictors, drop it as well and record the adjusted R-squared. Render.

```{r}
#| label: adjusted-r-squared-and-model-selection-5
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: adjusted-r-squared-and-model-selection-5-test
#| echo: true
tibble(dropped = c("verified_income", "debt_to_income", "credit_util",
                   "bankruptcy", "term", "credit_checks")) |>
  mutate(adj_r_squared = map_dbl(dropped, \(x) {
    lm(interest_rate ~ ., data = select(loans, -issue_month, -all_of(x))) |>
      glance() |>
      pull(adj.r.squared)
  })) |>
  arrange(desc(adj_r_squared)) |>
  mutate(adj_r_squared = num(adj_r_squared, digits = 4))
```

###

No row beats the new baseline of 0.2598, so backward elimination stops. "Predictors are eliminated one-at-a-time from the model until we cannot improve the model any further."

### Exercise 6

Replace the code in that chunk with the model that backward elimination chose, every predictor except `issue_month`, showing just the term and the estimate for each coefficient, to two decimal places. Render.

```{r}
#| label: adjusted-r-squared-and-model-selection-6
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: adjusted-r-squared-and-model-selection-6-test
#| echo: true
lm(interest_rate ~ . - issue_month, data = loans) |>
  tidy() |>
  select(term, estimate) |>
  mutate(estimate = num(estimate, digits = 2))
```

###

The remaining estimates barely moved from the full model's, because `issue_month` explained almost nothing. "It's common (and hip, at least in the statistical world) to refer to models that have undergone such predictor pruning as parsimonious."

### Exercise 7

Replace the code in that chunk so that it starts from the other end: for each of the seven predictors, fit a model with that predictor alone and record its adjusted R-squared. Show the results sorted from highest to lowest, to five decimal places. Render.

```{r}
#| label: adjusted-r-squared-and-model-selection-7
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: adjusted-r-squared-and-model-selection-7-test
#| echo: true
tibble(added = c("verified_income", "debt_to_income", "credit_util",
                 "bankruptcy", "term", "credit_checks", "issue_month")) |>
  mutate(adj_r_squared = map_dbl(added, \(x) {
    lm(interest_rate ~ ., data = select(loans, interest_rate, all_of(x))) |>
      glance() |>
      pull(adj.r.squared)
  })) |>
  arrange(desc(adj_r_squared)) |>
  mutate(adj_r_squared = num(adj_r_squared, digits = 5))
```

###

**Forward selection** adds `term` first. `issue_month` comes out below zero, worse than "a baseline model that has no predictors, which always has $R_{adj}^2 = 0$," although the chapter's list swaps it with `credit_checks` and gives `credit_checks` the negative value.

### Exercise 8

Replace the code in that chunk so that every model includes `term`, and each of the six other predictors is tried as a second predictor. Render.

```{r}
#| label: adjusted-r-squared-and-model-selection-8
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: adjusted-r-squared-and-model-selection-8-test
#| echo: true
tibble(added = c("verified_income", "debt_to_income", "credit_util",
                 "bankruptcy", "credit_checks", "issue_month")) |>
  mutate(adj_r_squared = map_dbl(added, \(x) {
    lm(interest_rate ~ ., data = select(loans, interest_rate, term, all_of(x))) |>
      glance() |>
      pull(adj.r.squared)
  })) |>
  arrange(desc(adj_r_squared)) |>
  mutate(adj_r_squared = num(adj_r_squared, digits = 5))
```

###

`credit_util` wins, lifting adjusted R-squared from 0.12855 to 0.20085. Continuing the same way adds `verified_income`, `credit_checks`, `debt_to_income`, and `bankruptcy`, then stops before `issue_month`: the same model backward elimination found.

### Exercise 9

Replace the code in that chunk with a table of the forward-selection path. Add the predictors one at a time in the order term, credit_util, verified_income, credit_checks, debt_to_income, bankruptcy, issue_month, and record the adjusted R-squared of the model after each addition, to four decimal places. Render.

```{r}
#| label: adjusted-r-squared-and-model-selection-9
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: adjusted-r-squared-and-model-selection-9-test
#| echo: true
predictors <- c("term", "credit_util", "verified_income", "credit_checks",
                "debt_to_income", "bankruptcy", "issue_month")

tibble(step = 1:7, added = predictors) |>
  mutate(adj_r_squared = map_dbl(step, \(i) {
    lm(interest_rate ~ ., data = select(loans, interest_rate, all_of(predictors[1:i]))) |>
      glance() |>
      pull(adj.r.squared)
  })) |>
  mutate(adj_r_squared = num(adj_r_squared, digits = 4))
```

###

The first two predictors lift adjusted R-squared to 0.20; the next four add only about 0.06 together. The chapter warns that "backward elimination and forward selection can arrive at different final models, because the decision for whether to include a given predictor or not depends on the other predictors that are already in the model."

### Exercise 10

Replace the table with a plot of adjusted R-squared at each step of the path, with the predictors on the x-axis in the order they were added. Render.

```{r}
#| label: adjusted-r-squared-and-model-selection-10
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: adjusted-r-squared-and-model-selection-10-test
#| echo: true
predictors <- c("term", "credit_util", "verified_income", "credit_checks",
                "debt_to_income", "bankruptcy", "issue_month")

tibble(step = 1:7, added = predictors) |>
  mutate(adj_r_squared = map_dbl(step, \(i) {
    lm(interest_rate ~ ., data = select(loans, interest_rate, all_of(predictors[1:i]))) |>
      glance() |>
      pull(adj.r.squared)
  })) |>
  ggplot(aes(x = fct_inorder(added), y = adj_r_squared, group = 1)) +
  geom_line() +
  geom_point()
```

###

The curve climbs steeply and then flattens. The drop at `issue_month`, from 0.2598 to 0.2597, is far too small to see at this scale.

### Exercise 11

Improve the plot. Label each point with its adjusted R-squared to four decimal places, and add a title, axis labels, a subtitle that states the main finding, and a caption giving the number of loans and the source. Render.

```{r}
#| label: adjusted-r-squared-and-model-selection-11
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: adjusted-r-squared-and-model-selection-11-test
#| echo: true
predictors <- c("term", "credit_util", "verified_income", "credit_checks",
                "debt_to_income", "bankruptcy", "issue_month")

tibble(step = 1:7, added = predictors) |>
  mutate(adj_r_squared = map_dbl(step, \(i) {
    lm(interest_rate ~ ., data = select(loans, interest_rate, all_of(predictors[1:i]))) |>
      glance() |>
      pull(adj.r.squared)
  })) |>
  ggplot(aes(x = fct_inorder(added), y = adj_r_squared, group = 1)) +
  geom_line() +
  geom_point() +
  geom_label(aes(label = round(adj_r_squared, 4)), vjust = -0.3, size = 3) +
  scale_y_continuous(limits = c(0.1, 0.3)) +
  labs(title = "Forward Selection for Loan Interest Rates",
       subtitle = "Term and credit utilization do most of the work; adding issue month lowers adjusted R-squared",
       x = "Predictor Added",
       y = "Adjusted R-squared",
       caption = "Up to 10,000 Lending Club loans, January to March 2018; openintro package")
```

###

"It is highly advised that *before* you begin the model selection process, you decide what a "meaningful" difference in adjusted $R^2$ is for the context of your data." With a threshold of even 0.01, forward selection would have stopped after `credit_checks`, at four predictors instead of six.

### Exercise 12

Commit `analysis.qmd` with the message `"Select a model with adjusted R-squared"` and push.

```{r}
#| label: adjusted-r-squared-and-model-selection-12
#| echo: false
learnr2::question("In the bash Terminal, run git log --oneline -3. CP/CR.", type = "reflection")
```

###

```
a41d7e9 (HEAD -> main, origin/main) Select a model with adjusted R-squared
3f8a1c2 Fit models with categorical and many predictors
9be04d2 Initial commit
```

###

Adjusted R-squared is only one yardstick. "In fact, many statisticians discourage the use of stepwise regression *alone* for model selection and advocate, instead, for a more thoughtful approach that carefully considers the research focus and features of the data."

## Summary

This tutorial covered [Chapter 8: Linear regression with multiple predictors](https://openintrostat.github.io/ims/model-mlr.html) from [*Introduction to Modern Statistics*](https://openintrostat.github.io/ims/) by Mine Çetinkaya-Rundel and Johanna Hardin. You have learned how a regression model uses indicator and categorical predictors, how to fit and read a model with many predictors at once, and how to choose among models with adjusted R-squared, using backward elimination and forward selection.

For what these same models say about the population the loans came from, read [Chapter 25: Inference for regression with multiple predictors](https://openintrostat.github.io/ims/inf-model-mlr.html), which returns to the `loans` data. For more on turning models into tidy tables, see the **broom** [introduction](https://broom.tidymodels.org/articles/broom.html).

### Exercise 1

Render one last time.

```{r}
#| label: summary-1
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd", start = 0). CP/CR.', type = "reflection")
```

###

<pre><code>---
title: "Linear Regression with Multiple Predictors"
author: "Your Name"
format: html
execute:
  echo: false
---

&#96;&#96;&#96;{r}
#| message: false
library(tidyverse)
library(openintro)
library(broom)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| cache: true
loans <- loans_full_schema |>
  mutate(credit_util = total_credit_utilized / total_credit_limit,
         bankruptcy = as.factor(if_else(public_record_bankrupt == 0, 0, 1)),
         verified_income = droplevels(verified_income)) |>
  rename(credit_checks = inquiries_last_12m) |>
  select(interest_rate, verified_income, debt_to_income, credit_util,
         bankruptcy, term, credit_checks, issue_month)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
lm(interest_rate ~ credit_util + verified_income, data = loans) |>
  augment() |>
  ggplot(aes(x = credit_util, color = verified_income)) +
  geom_point(aes(y = interest_rate), alpha = 0.1) +
  geom_line(aes(y = .fitted), linewidth = 1) +
  scale_x_continuous(labels = scales::label_percent()) +
  labs(title = "Interest Rate by Credit Utilization and Income Verification",
       subtitle = "At every utilization, verified borrowers are predicted to pay 3.3 points more",
       x = "Credit Utilization",
       y = "Interest Rate (%)",
       color = "Income Verification",
       caption = "9,998 Lending Club loans, January to March 2018; openintro package") +
  theme(legend.position = "bottom")
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
predictors <- c("term", "credit_util", "verified_income", "credit_checks",
                "debt_to_income", "bankruptcy", "issue_month")

tibble(step = 1:7, added = predictors) |>
  mutate(adj_r_squared = map_dbl(step, \(i) {
    lm(interest_rate ~ ., data = select(loans, interest_rate, all_of(predictors[1:i]))) |>
      glance() |>
      pull(adj.r.squared)
  })) |>
  ggplot(aes(x = fct_inorder(added), y = adj_r_squared, group = 1)) +
  geom_line() +
  geom_point() +
  geom_label(aes(label = round(adj_r_squared, 4)), vjust = -0.3, size = 3) +
  scale_y_continuous(limits = c(0.1, 0.3)) +
  labs(title = "Forward Selection for Loan Interest Rates",
       subtitle = "Term and credit utilization do most of the work; adding issue month lowers adjusted R-squared",
       x = "Predictor Added",
       y = "Adjusted R-squared",
       caption = "Up to 10,000 Lending Club loans, January to March 2018; openintro package")
&#96;&#96;&#96;
</code></pre>

###

The page holds the chapter's two halves: a model with several kinds of predictors, and a way to decide which predictors to keep. "There is often no perfect or "right" final model, however, using the adjusted $R^2$ value is one way to identify important predictor variables for a final regression model."

### Exercise 2

Publish your analysis to GitHub Pages. In a bash Terminal, run:

```
quarto publish gh-pages analysis.qmd
```

```{r}
#| label: summary-2
#| echo: false
learnr2::question("Paste the URL of your published page.", type = "reflection")
```

###

Your answer should look like `https://your-github-username.github.io/linear-regression-with-multiple-predictors/`.

###

"Model building is an extensive topic, and we scratch the surface here by defining and utilizing the adjusted $R^2$ value." Other yardsticks for the same job, such as AIC and BIC, come in later courses.

### Exercise 3

Commit and push any remaining changes.

```{r}
#| label: summary-3
#| echo: false
learnr2::question("Paste the URL of your GitHub repo.", type = "reflection")
```

###

Your answer should look like `https://github.com/your-github-username/linear-regression-with-multiple-predictors`.

###

Every number in this tutorial describes these 10,000 loans and nothing more. "Note that these values are sample statistics and in the case where the observed data is a random sample from a target population that we are interested in making inferences about, they are estimates of the population parameters."

## Your Answers

```{r}
#| label: your-answers-1
#| echo: false
learnr2::question(
  "How many minutes, approximately, did it take you to complete this
  tutorial? For example, an hour and a half would be 90 minutes.",
  type = "reflection_editable",
  validate = "integer"
)
```

```{r}
#| label: your-answers-2
#| echo: false
learnr2::download_answers_button(filename_prefix = "linear-regression-multiple")
```
