---
title: "Hello Data"
subtitle: "Cases, variables, associations, and the experiment that showed stents cause strokes."
format: live-html
engine: knitr
toc: true
toc-depth: 2
---

{{< include _extensions/r-wasm/live/_knitr.qmd >}}

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(openintro)
library(usdata)
knitr::opts_chunk$set(out.width = "90%")
```

## Student Information

```{r}
#| label: student-information-1
#| echo: false
learnr2::student_info()
```

## Introduction

This tutorial covers [Chapter 1: Hello data](https://openintrostat.github.io/ims/data-hello.html) from [*Introduction to Modern Statistics*](https://openintrostat.github.io/ims/) by Mine Çetinkaya-Rundel and Johanna Hardin. You will learn the vocabulary of data (cases, variables, and their types), how to spot an association between two variables, and why a randomized experiment can show cause and effect when an observational study cannot.

### Exercise 1

You should be doing this tutorial in a repo named `hello-data`. If you are not, create one and then connect to it. You may need to restart this tutorial after you do so.

Create a new file, `analysis.qmd`, with the title `"Hello Data"` and your name as the author. In a bash Terminal, render it:

```
quarto render analysis.qmd
```

Open `analysis.html` with Live Server (right-click it in the Explorer → **Open with Live Server**) and keep the tab open. It refreshes on every render. Going forward, we will just tell you to "Render" when we want you to take these steps.

Create a `.gitignore` with `analysis_files` followed by a blank line. Commit and push. Use AI however you like.

If `show_file()` fails, it is probably because you have not yet loaded `library(learnr2)` in the R Terminal.

```{r}
#| label: introduction-1
#| echo: false
learnr2::question('In the R Terminal, run show_file(".gitignore"). CP/CR.', type = "reflection")
```

###

```
analysis_files
```

###

"Statistics is the study of how best to collect, analyze, and draw conclusions from data." This chapter is about the first step: what data looks like and how it is collected.

### Exercise 2

Add a chunk to `analysis.qmd` that loads the **[tidyverse](https://www.tidyverse.org/)**, with `#| message: false` so the startup messages stay out of the rendered page. Add `execute: echo: false` to the YAML header. Render.

```{r}
#| label: introduction-2
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

<pre><code>#| message: false
library(tidyverse)
</code></pre>

###

The **tidyverse** is a family of packages for importing, reshaping, summarizing, and plotting data frames. It is named for *tidy data*, the way of organizing data that this chapter recommends.

### Exercise 3

Add **[openintro](https://openintrostat.github.io/openintro/)** to that chunk. Render.

```{r}
#| label: introduction-3
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

<pre><code>#| message: false
library(tidyverse)
library(openintro)
</code></pre>

###

**openintro** holds 246 datasets, nearly every one used in the book. Two of them, `stent30` and `stent365`, come from the medical experiment this chapter opens with.

### Exercise 4

Add **[usdata](https://openintrostat.github.io/usdata/)** to that chunk. Render.

```{r}
#| label: introduction-4
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

<pre><code>#| message: false
library(tidyverse)
library(openintro)
library(usdata)
</code></pre>

###

**usdata** holds demographic data on US states and counties. Its `county` data frame describes all 3,142 counties in the country.

## Case study: using stents to prevent strokes

Stents are small mesh tubes placed inside narrow or weak arteries. Researchers enrolled 451 patients at risk of stroke in an experiment to answer one question: "Does the use of stents reduce the risk of stroke?"

### Exercise 1

Add a chunk that prints `stent365`. Render.

```{r}
#| label: case-study-using-stents-to-prevent-strokes-1
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: case-study-using-stents-to-prevent-strokes-1-test
#| echo: true
stent365
```

###

Each row is one patient: `group` records whether they received a stent, and `outcome` whether they had a stroke within 365 days. Every row we can see is a stroke in the treatment group, because the data happens to be sorted that way.

### Exercise 2

```{r}
#| label: case-study-using-stents-to-prevent-strokes-2
#| echo: false
learnr2::question("In the R Terminal, run ?stent365. Copy and paste the Description and the Source sections.", type = "reflection")
```

###

```
Description:

     An experiment that studies effectiveness of stents in treating
     patients at risk of stroke with some unexpected results. 'stent30'
     represents the results 30 days after stroke and 'stent365'
     represents the results 365 days after stroke.

Source:

     Chimowitz MI, Lynn MJ, Derdeyn CP, et al. 2011. Stenting versus
     Aggressive Medical Therapy for Intracranial Arterial Stenosis.
     New England Journal of Medicine 365:993-1003.
```

###

Read the description closely: the results are measured 30 and 365 days after each patient *enrolled* in the study, not after a stroke, as the book makes clear. Help pages are written by people, so check them against the original source when a detail matters.

### Exercise 3

Replace the code in that chunk so that it counts the number of patients in each `group`. Render.

```{r}
#| label: case-study-using-stents-to-prevent-strokes-3
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: case-study-using-stents-to-prevent-strokes-3-test
#| echo: true
stent365 |>
  count(group)
```

###

Researchers randomly assigned 224 patients to the treatment group, who received a stent plus medical management, and 227 to the control group, who received the same medical management without a stent. "Random assignment organizes the participants in a study into groups that are roughly equal on all aspects," which is what allows an experiment to show cause and effect.

### Exercise 4

Replace the code in that chunk so that it counts patients by both `group` and `outcome`. Render.

```{r}
#| label: case-study-using-stents-to-prevent-strokes-4
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: case-study-using-stents-to-prevent-strokes-4-test
#| echo: true
stent365 |>
  count(group, outcome)
```

###

45 of the 224 treatment patients had a stroke within a year, against 28 of the 227 control patients, the same counts as the book's Table 1.2. The control group "provides a reference point against which we can measure the medical impact of stents."

### Exercise 5

Replace the code in that chunk so that it calculates, for each `group`, the proportion of patients who had a stroke. Render.

```{r}
#| label: case-study-using-stents-to-prevent-strokes-5
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: case-study-using-stents-to-prevent-strokes-5-test
#| echo: true
stent365 |>
  summarize(prop_stroke = mean(outcome == "stroke"),
            .by = group)
```

###

"A summary statistic is a single number summarizing data from a sample." These two, 20% for the treatment group and 12% for the control group, deliver a surprise: "an additional 8% of patients in the treatment group had a stroke!"

### Exercise 6

The study also recorded outcomes 30 days after enrollment, in `stent30`. Start a new chunk that stacks `stent30` and `stent365` into one tibble named `stent`, with a `time` column recording `"30 days"` or `"365 days"`. In the same chunk, calculate the proportion of patients who had a stroke for each combination of `time` and `group`. Render.

```{r}
#| label: case-study-using-stents-to-prevent-strokes-6
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: case-study-using-stents-to-prevent-strokes-6-test
#| echo: true
stent <- bind_rows(`30 days` = stent30, `365 days` = stent365, .id = "time")

stent |>
  summarize(prop_stroke = mean(outcome == "stroke"),
            .by = c(time, group))
```

###

The gap opened early: within 30 days, 14.7% of the treatment group had a stroke, against 5.7% of the control group. Both proportions grew by 365 days, but the treatment group stayed about 8 points worse.

### Exercise 7

Replace the summary in that chunk with a bar chart of the proportion of patients who had a stroke, by `group`, with one panel for each `time`. Render.

```{r}
#| label: case-study-using-stents-to-prevent-strokes-7
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: case-study-using-stents-to-prevent-strokes-7-test
#| echo: true
stent |>
  summarize(prop_stroke = mean(outcome == "stroke"),
            .by = c(time, group)) |>
  ggplot(aes(x = group, y = prop_stroke)) +
  geom_col() +
  facet_wrap(~ time)
```

###

The treatment bar is taller in both panels, the opposite of what doctors expected. The published analysis found "compelling evidence of harm by stents in this study."

### Exercise 8

Improve the bar chart. Put the 30-day panel first, label each bar with its percentage, and show the y-axis as percentages. Add a title, axis labels, a subtitle that states the main finding, and a caption giving the number of patients and the source. Render.

```{r}
#| label: case-study-using-stents-to-prevent-strokes-8
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: case-study-using-stents-to-prevent-strokes-8-test
#| echo: true
stent |>
  summarize(prop_stroke = mean(outcome == "stroke"),
            .by = c(time, group)) |>
  mutate(time = fct_relevel(time, "30 days")) |>
  ggplot(aes(x = group, y = prop_stroke)) +
  geom_col() +
  geom_text(aes(label = scales::label_percent(accuracy = 1)(prop_stroke)),
            vjust = -0.5) +
  facet_wrap(~ time) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(title = "Stroke Rates in the Stent Experiment",
       subtitle = "Patients who received a stent had more strokes, not fewer",
       x = "Group",
       y = "Patients Who Had a Stroke",
       caption = "451 patients; Chimowitz et al. (2011)")
```

###

"Do not generalize the results of this study to all patients and all stents." The patients were volunteers with specific characteristics, and the study tested only one stent, the self-expanding Wingspan.

### Exercise 9

Commit `analysis.qmd` with the message `"Analyze the stent experiment"` and push.

```{r}
#| label: case-study-using-stents-to-prevent-strokes-9
#| echo: false
learnr2::question("In the bash Terminal, run git log --oneline -3. CP/CR.", type = "reflection")
```

###

```
3f2c1a7 (HEAD -> main, origin/main) Analyze the stent experiment
9be04d2 Initial commit
```

###

Even if stents did nothing, the two groups would rarely have exactly equal stroke rates, just as 100 coin flips rarely give exactly 50 heads. So the real question is, "if in fact stents have no effect, how likely is it that we observe such a large difference?"

## Data basics

"Effective presentation and description of data is a first step in most analyses." The book builds its vocabulary for data on two data frames: a sample of 50 loans and every county in the United States.

### Exercise 1

Start a new chunk that prints `loan50`. Render.

```{r}
#| label: data-basics-1
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-1-test
#| echo: true
loan50
```

###

"Each row in the table represents a single loan," made through Lending Club, a peer-to-peer lending company. With 18 columns, the tibble cannot fit them all across the page, so it lists the rest by name below the table.

### Exercise 2

```{r}
#| label: data-basics-2
#| echo: false
learnr2::question("In the R Terminal, run ?loan50. Copy and paste the Description and the first line of the Format section.", type = "reflection")
```

###

```
Description:

     This dataset represents thousands of loans made through the
     Lending Club platform, which is a platform that allows individuals
     to lend to other individuals. [...] It is important to keep that
     last part in mind, since this dataset only represents loans
     actually made, i.e. do not mistake this data for loan
     applications!

Format:

     A data frame with 10,000 observations on the following 55
     variables.
```

###

This page also documents `loans_full_schema`, with 10,000 loans and 55 variables; `loan50` is a sample of 50 of those loans, with 18 of the variables. Its warning matters: riskier applicants who never received a loan do not appear in the data at all.

### Exercise 3

Replace the code in that chunk so that it prints the names of all 18 variables. Render.

```{r}
#| label: data-basics-3
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-3-test
#| echo: true
names(loan50)
```

###

"The formal name for a row is a case or observation or unit of observation," and each column "is referred to as a variable." The book works with only 7 of these 18 variables.

### Exercise 4

Replace the code in that chunk so that it keeps only `loan_amount`, `interest_rate`, `term`, `grade`, `state`, `total_income`, and `homeownership`, and shows the type and first few values of each. Render.

```{r}
#| label: data-basics-4
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-4-test
#| echo: true
loan50 |>
  select(loan_amount, interest_rate, term, grade, state, total_income, homeownership) |>
  glimpse()
```

###

The first loan is for $22,000 at 10.90% interest, to a borrower in New Jersey with an income of $59,000. A data frame where each row is a case, each column is a variable, and each cell is a single value "is commonly referred to as tidy data," a term from Hadley Wickham's paper "[Tidy Data](https://doi.org/10.18637/jss.v059.i10)."

### Exercise 5

Replace the code in that chunk so that it calculates the mean and the median of `interest_rate`. Render.

```{r}
#| label: data-basics-5
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-5-test
#| echo: true
loan50 |>
  summarize(mean_rate = mean(interest_rate),
            median_rate = median(interest_rate))
```

###

A variable is *numerical* when "it is sensible to add, subtract, or take averages with those values," as it is for these interest rates. Because a rate can take any value in a range, like 9.92 or 10.90, `interest_rate` is *continuous* rather than *discrete*.

### Exercise 6

Replace the code in that chunk so that it counts the loans from each `state`, from most to fewest. Render.

```{r}
#| label: data-basics-6
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-6-test
#| echo: true
loan50 |>
  count(state, sort = TRUE)
```

###

You can count states but not average them, which makes `state` a *categorical* variable whose possible values are its *levels*. The book's example is a telephone area code: it is a number, but "the average, sum, and difference of area codes does not have any clear meaning."

### Exercise 7

Replace the code in that chunk so that it counts the loans in each `grade`. Render.

```{r}
#| label: data-basics-7
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-7-test
#| echo: true
loan50 |>
  count(grade)
```

###

Grade B is the most common, with 19 loans. The book says grades run from A through G and represent "the quality of the loan and its likelihood of being repaid," yet no loan here is graded F or G.

### Exercise 8

Replace the code in that chunk so that it prints the levels of `grade`: every value it is allowed to take, not just the ones that appear. Render.

```{r}
#| label: data-basics-8
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-8-test
#| echo: true
levels(loan50$grade)
```

###

`grade` has eight levels: A through G, plus an empty one, `""`, that no loan uses. Its levels have a natural order, which makes it an *ordinal* variable, while a categorical variable without an order, like `state`, is *nominal*.

### Exercise 9

Replace the code in that chunk so that it counts the loans with each `term`. Render.

```{r}
#| label: data-basics-9
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-9-test
#| echo: true
loan50 |>
  count(term)
```

###

`term` is "always set as a whole number of months," which makes it *discrete*, but only two values appear: 36 loans run 36 months and 14 run 60. A numerical variable with so few values is often analyzed as if it were categorical.

### Exercise 10

Start a new chunk that prints `county`. Render.

```{r}
#| label: data-basics-10
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-10-test
#| echo: true
county
```

###

Each of the 3,142 rows is a county, described by 15 variables covering population, poverty, housing, income, and education. Unlike `loan50`, this data frame is not a sample: it includes every county in the country.

### Exercise 11

```{r}
#| label: data-basics-11
#| echo: false
learnr2::question("In the R Terminal, run ?county. Copy and paste the first line of the Format section, the entry for smoking_ban, and the Source section.", type = "reflection")
```

###

```
Format:

     A data frame with 3142 observations on the following 14 variables.

     smoking_ban Describes whether the type of county-level smoking ban
          in place in 2010, taking one of the values '"none"',
          '"partial"', or '"comprehensive"'.

Source:

     These data were collected from Census Quick Facts (no longer
     available as of 2020) and its accompanying pages. Smoking ban data
     were from a variety of sources.
```

###

The page says 14 variables, but it documents 15, as many as the data frame has. "It is always important to be sure we know what each variable means and its units of measurement," and sometimes that means noticing where the documentation is wrong.

### Exercise 12

Replace the code in that chunk so that it counts the missing values in each column of `county`, with one row per column. Render.

```{r}
#| label: data-basics-12
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-12-test
#| echo: true
county |>
  summarize(across(everything(), \(x) sum(is.na(x)))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "n_missing")
```

###

No column is missing more than three values except `smoking_ban`, which is missing 580. The help page warned that the smoking ban data "were from a variety of sources."

### Exercise 13

Replace the code in that chunk so that it counts the counties with each value of `smoking_ban`. Render.

```{r}
#| label: data-basics-13
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-13-test
#| echo: true
county |>
  count(smoking_ban)
```

###

The help page promises none, partial, or comprehensive, yet not one county is comprehensive and 580 are missing. Before using this variable, you would need to find out whether those 580 are the counties with comprehensive bans.

### Exercise 14

Replace the code in that chunk so that it counts the counties with each value of `median_edu`. Render.

```{r}
#| label: data-basics-14
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-14-test
#| echo: true
county |>
  count(median_edu)
```

###

`median_edu` "is a categorical variable, but the levels have a natural ordering," running from below high school to a bachelor's degree, so it is ordinal. In 3,092 of the 3,142 counties, the median resident finished high school but not college.

### Exercise 15

Replace the code in that chunk so that it keeps `name`, `state`, `homeownership`, `multi_unit`, `pop_change`, and `median_hh_income`, drops every county with a missing value in any of them, and stores the result as `county_clean`. Print `county_clean`. Render.

```{r}
#| label: data-basics-15
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-15-test
#| echo: true
county_clean <- county |>
  select(name, state, homeownership, multi_unit, pop_change, median_hh_income) |>
  drop_na()

county_clean
```

###

`county_clean` has 3,137 counties, 5 fewer than `county`. Three Alaska areas have no 2017 population, so their population change cannot be computed.

### Exercise 16

Delete the line that prints `county_clean`, so that the chunk only creates it. Render.

```{r}
#| label: data-basics-16
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

<pre><code>county_clean <- county |>
  select(name, state, homeownership, multi_unit, pop_change, median_hh_income) |>
  drop_na()
</code></pre>

###

`homeownership` is the percent of homes occupied by their owners, and `multi_unit` is the percent of housing units in structures like apartments and condos. Both were measured from 2006 to 2010.

### Exercise 17

Add `#| cache: true` to that chunk. Render.

```{r}
#| label: data-basics-17
#| echo: false
learnr2::question("In the bash Terminal, run ls. CP/CR.", type = "reflection")
```

###

```
analysis.html  analysis.qmd  analysis_cache  analysis_files
```

###

`analysis_cache` now holds the saved `county_clean`, so later renders skip rebuilding it. `pop_change` is the percent change in population from 2010 to 2017, and `median_hh_income` is in US dollars.

### Exercise 18

Your Source Control count just jumped, because Git sees every file in the new cache directory. Add `analysis_cache` to `.gitignore`, and the count drops back.

```{r}
#| label: data-basics-18
#| echo: false
learnr2::question('In the R Terminal, run show_file(".gitignore"). CP/CR.', type = "reflection")
```

###

```
analysis_files
analysis_cache
```

###

Each row of `county` describes a whole county. A relationship you find in this data holds between counties, not necessarily between the people who live in them.

### Exercise 19

Start a new chunk with a scatterplot of `homeownership` against `multi_unit`, using `county_clean`. Add a title, axis labels, a subtitle that states the main finding, and a caption giving the number of counties. Render.

```{r}
#| label: data-basics-19
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-19-test
#| echo: true
county_clean |>
  ggplot(aes(x = multi_unit, y = homeownership)) +
  geom_point(alpha = 0.3) +
  labs(title = "Homeownership and Multi-Unit Housing in US Counties",
       subtitle = "Counties with more apartments and condos have lower homeownership",
       x = "Housing Units in Multi-Unit Structures (%)",
       y = "Homeownership Rate (%)",
       caption = "3,137 US counties")
```

###

Counties with more multi-unit housing tend to have lower homeownership, a *negative association*. When two variables show some connection with one another, "they are called associated variables."

### Exercise 20

One county sits alone at 0% homeownership. Start a new chunk that shows the three counties in `county_clean` with the lowest homeownership rates, along with their `state`, `homeownership`, and `multi_unit`. Render.

```{r}
#| label: data-basics-20
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-20-test
#| echo: true
county_clean |>
  slice_min(homeownership, n = 3) |>
  select(name, state, homeownership, multi_unit)
```

###

Kalawao County, Hawaii, has 88 residents: it is a former settlement for people with Hansen's disease, now a national historical park, where no one owns their home. The other two, the Bronx and Manhattan, fit the trend, with nearly all their housing in multi-unit buildings.

### Exercise 21

Replace the code in that chunk with a scatterplot of `pop_change` against `median_hh_income`, using `county_clean`. Render.

```{r}
#| label: data-basics-21
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-21-test
#| echo: true
county_clean |>
  ggplot(aes(x = median_hh_income, y = pop_change)) +
  geom_point(alpha = 0.3)
```

###

Counties with higher median household incomes tend to have grown faster, a *positive association*, though many counties break the trend. Two variables with no evident relationship are called *independent*: "Associated or independent, not both."

### Exercise 22

Improve the scatterplot. Add a trend line, show income in dollars, and add a title, axis labels, a subtitle that states the main finding, and a caption giving the number of counties. Render.

```{r}
#| label: data-basics-22
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd"). CP/CR.', type = "reflection")
```

###

```{r}
#| label: data-basics-22-test
#| echo: true
county_clean |>
  ggplot(aes(x = median_hh_income, y = pop_change)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE) +
  scale_x_continuous(labels = scales::label_dollar()) +
  labs(title = "Population Change and Income in US Counties, 2010 to 2017",
       subtitle = "Counties with higher median household incomes tended to grow faster",
       x = "Median Household Income",
       y = "Population Change (%)",
       caption = "3,137 US counties")
```

###

Income sits on the x-axis because, if we suspect it drives growth, it is the *explanatory variable* and population change is the *response variable*. But "the act of labeling the variables in this way does nothing to guarantee that a causal relationship exists."

### Exercise 23

Commit `analysis.qmd` and `.gitignore` with the message `"Explore loans and counties"` and push.

```{r}
#| label: data-basics-23
#| echo: false
learnr2::question("In the bash Terminal, run git log --oneline -3. CP/CR.", type = "reflection")
```

###

```
8d41e0b (HEAD -> main, origin/main) Explore loans and counties
3f2c1a7 Analyze the stent experiment
9be04d2 Initial commit
```

###

The county data is an *observational study*: researchers "merely observe the data that arise," so it can show associations but "cannot by themselves show a causal connection." The stent study could show that stents cause harm only because it was a randomized experiment.

## Summary

This tutorial covered [Chapter 1: Hello data](https://openintrostat.github.io/ims/data-hello.html) from [*Introduction to Modern Statistics*](https://openintrostat.github.io/ims/) by Mine Çetinkaya-Rundel and Johanna Hardin. You have learned the vocabulary of data (cases, variables, and their types), how to spot an association between two variables, and why a randomized experiment can show cause and effect when an observational study cannot.

For more on organizing data, read Hadley Wickham's "[Tidy Data](https://doi.org/10.18637/jss.v059.i10)." For the stent study itself, read the original [article](https://doi.org/10.1056/NEJMoa1105335) in the *New England Journal of Medicine*.

### Exercise 1

Render one last time.

```{r}
#| label: summary-1
#| echo: false
learnr2::question('In the R Terminal, run show_file("analysis.qmd", start = 0). CP/CR.', type = "reflection")
```

###

<pre><code>---
title: "Hello Data"
author: "Your Name"
format: html
execute:
  echo: false
---

&#96;&#96;&#96;{r}
#| message: false
library(tidyverse)
library(openintro)
library(usdata)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
stent365 |>
  summarize(prop_stroke = mean(outcome == "stroke"),
            .by = group)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
stent <- bind_rows(`30 days` = stent30, `365 days` = stent365, .id = "time")

stent |>
  summarize(prop_stroke = mean(outcome == "stroke"),
            .by = c(time, group)) |>
  mutate(time = fct_relevel(time, "30 days")) |>
  ggplot(aes(x = group, y = prop_stroke)) +
  geom_col() +
  geom_text(aes(label = scales::label_percent(accuracy = 1)(prop_stroke)),
            vjust = -0.5) +
  facet_wrap(~ time) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(title = "Stroke Rates in the Stent Experiment",
       subtitle = "Patients who received a stent had more strokes, not fewer",
       x = "Group",
       y = "Patients Who Had a Stroke",
       caption = "451 patients; Chimowitz et al. (2011)")
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
loan50 |>
  count(term)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
county |>
  count(median_edu)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| cache: true
county_clean <- county |>
  select(name, state, homeownership, multi_unit, pop_change, median_hh_income) |>
  drop_na()
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
county_clean |>
  ggplot(aes(x = multi_unit, y = homeownership)) +
  geom_point(alpha = 0.3) +
  labs(title = "Homeownership and Multi-Unit Housing in US Counties",
       subtitle = "Counties with more apartments and condos have lower homeownership",
       x = "Housing Units in Multi-Unit Structures (%)",
       y = "Homeownership Rate (%)",
       caption = "3,137 US counties")
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
county_clean |>
  ggplot(aes(x = median_hh_income, y = pop_change)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE) +
  scale_x_continuous(labels = scales::label_dollar()) +
  labs(title = "Population Change and Income in US Counties, 2010 to 2017",
       subtitle = "Counties with higher median household incomes tended to grow faster",
       x = "Median Household Income",
       y = "Population Change (%)",
       caption = "3,137 US counties")
&#96;&#96;&#96;
</code></pre>

###

The page holds the chapter in miniature: an experiment that supports a causal claim, and an observational study that shows two associations but cannot. Data "can be organized in many ways but tidy data … lends itself most easily to statistical analysis."

### Exercise 2

Publish your analysis to GitHub Pages. In a bash Terminal, run:

```
quarto publish gh-pages analysis.qmd
```

```{r}
#| label: summary-2
#| echo: false
learnr2::question("Paste the URL of your published page.", type = "reflection")
```

###

Your answer should look like `https://your-github-username.github.io/hello-data/`.

###

The stent study appeared in the *New England Journal of Medicine* in 2011. Its central finding, more strokes with stents than without, fits in the two panels on your page.

### Exercise 3

Commit and push any remaining changes.

```{r}
#| label: summary-3
#| echo: false
learnr2::question("Paste the URL of your GitHub repo.", type = "reflection")
```

###

Your answer should look like `https://github.com/your-github-username/hello-data`.

###

"In general, association does not imply causation." Of every idea in this chapter, that is the one to carry into the rest of the book.

## Your Answers

```{r}
#| label: your-answers-1
#| echo: false
learnr2::question(
  "How many minutes, approximately, did it take you to complete this
  tutorial? For example, an hour and a half would be 90 minutes.",
  type = "reflection_editable",
  validate = "integer"
)
```

```{r}
#| label: your-answers-2
#| echo: false
learnr2::download_answers_button(filename_prefix = "01-hello-data")
```
